---
title: "SI_Project_P1"
author: "Rahul Jain"
date: "30/01/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

The main purpose of this project is to investigate the exponential distribution and then compare it to the CLT.
For this analysis we will use lambda as 0.2 and simulate the distribution of averages of 40 exponentials over 
1000 simulations.

## Library Import

```{r echo=TRUE}
library(ggplot2)
```

## Simulations

Simulate the data from exponential distribution

```{r echo=TRUE}
set.seed(42)
lambda = 0.2
nosim = 1000
count = 40

expData <- matrix(sample(rexp(nosim*count,lambda)),nosim,count)
expMns <- apply(expData,1,mean)
```

## Sample Mean Comparison with theoretical Mean

#### Sample Mean Calculation

```{r echo=TRUE}
paste("Mean of sample means from simulation = ",mean(expMns))
```

#### Theoretical Mean Calculation

```{r echo=TRUE}
paste("Theoretical Mean = ",1/lambda)
```

#### Comparison

There is only a small difference between theoretical mean and sample mean

```{r echo=TRUE}
1/lambda - mean(expMns)
```


## Sample Variance Comparison with theoretical variance

#### Sample Variance Calculation

```{r echo=TRUE}
paste("Variance of Sample means from simulation = ",var(expMns))
```

#### Theoretical Variance Calculation

```{r echo=TRUE}
paste("Theoretical Variance = ",1/((lambda*sqrt(count))^2))
```

#### Comparison

There is very little difference between theoretical variance and variance of simulated means

```{r echo=TRUE}
1/((lambda*sqrt(count))^2) - var(expMns)
```

## Distribution

Distribution of Sample Means from exponential distribution

```{r echo=TRUE}
xfit <- seq(min(expMns), max(expMns), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(count)))
hist(expMns,breaks=count,prob=T,col = "green",xlab = "Sample Means",main = "Distribution of Sample Means",ylab="Density")
lines(xfit,yfit,col="black",lwd=2)
```

#### Comparison with normal distribution

Quantile plots show that the distribution of Sample Means from exponential distribution is 
approximately normal

```{r echo=TRUE}
qqnorm(expMns)
qqline(expMns,col="red",lwd=2.5)
```